Modelling the Real Time State of a Public Transport Road Network

### 2017-04-21 realtime-filter/realtime-filter.Rnw
Type up eqns

%\documentclass{IEEEtran}
\documentclass[draftcls,a4paper,onecolumn]{IEEEtran}

@@

So, depending on the data, we apply different likelihood functions.

\subsubsection[L1]{$L_1$: not at a stop or intersection}

In situations where the bus is neither at a bus stop nor at an intersection,
the likelihood is simply determined from the distance between the particle and 
vehicle's reported position.
The particle filter framework makes this easy,
as each particle has a single value of distance into trip, $d_k^{(i)}$,
with which we can compute the latitude and longitude values based on 
the routes shape information, $z_k^{(i)} = h(d_k{(i)})$.

\subsubsection[L1b]{$L_1'$: at stop}

If the particle is considered to be at a stop, then we apply the following likelihood.
Define a parameter $\varepsilon_S$ as a boundary around the stop;
if observation is within that boundary likelihood is 1, else 0.

\subsubsection[L1c]{$L_1''$: at intersection}

If the particle is considered to be at an intersection, we apply the following likelihood.
Similar to $L_1'$, however need to check observation is heading into the intersection
(and not out of it) by using two boundaries around the intersection and a point $\varepsilon_I$ away.

Then 
\begin{equation}
  L_1(\bX_k^{(i)} | \bY_k) =
  \begin{cases}
    \frac{1}{2\pi\sigma_y} e^{\frac{\|g(h(d_k^{(i)})|\bY_k)\|^2}{2\sigma_y^2}} & \text{ if particle neither at stop nor intersection } \\
    \frac{1}{c} \mathbbm{1}_{d(h(S_{s_k{[i]}}^d), Y_k) < \varepsilon_S}    & \text{ if particle at stop } \\
    \frac{1}{c} \mathbbm{1}_{d(h(R_{r_k{[i]}}^d), Y_k) < \varepsilon_I \cap d(h(R_{r_k{[i]}}^d-\varepsilon_I), Y_k) < \varepsilon_I}    & \text{ if particle at intersection }
  \end{cases}
\end{equation}

\subsubsection[L2]{$L_2$: not at a stop}

If the particle is not at a stop, we determine a likelihood for the arrival and departure
times at the last stop, if they are available.

First check that the most recent stop time update agrees with the particle;
if the stop time update is for stop $j$, and the particles state has $s_k^{(i)} < j$,
then the particle is not plausible;
otherwise use normal distribution for arrival and departure times,
or whichever are available.

\subsubsection[L2b]{$L_2'$: at a stop}

If particle is at a stop, again determine whether or not it agrees with data.
If it does, then check data doesn't imply particle has departed,
then just use likelihood on arrival time.

>>>>

\documentclass[9pt]{IEEEtran}
%\documentclass[draftcls,a4paper,onecolumn]{IEEEtran}

@@

So, depending on the particles' states, we apply different likelihood functions.

\begin{enumerate}
%\subsubsection[L1]{$L_1$: not at a stop or intersection}
\item
In situations where the bus is neither at a bus stop nor at an intersection,
the likelihood is simply determined from the distance between the particle and 
vehicle's reported position.
The particle filter framework makes this easy,
as each particle has a single value of distance into trip, $d_k^{(i)}$,
with which we can compute the latitude and longitude values based on 
the routes shape information, $z_k^{(i)} = h(d_k{(i)})$.

\begin{equation}
  \label{eq:l1}
  \log L_1 (\bX_k^{(i)} | \bY_k) = - \log(2\pi\sigma_y) - \frac{\|g(h(d_k^{(i)})|\bY_k)\|^2}{2\sigma_y^2}
\end{equation}

%\subsubsection[L1b]{$L_1'$: at stop}
\item
If the particle is considered to be at a stop, then we apply the following likelihood.
Define a parameter $\varepsilon_S$ as a boundary around the stop;
if observation is within that boundary likelihood is 1, else 0.
Let $B_s$ be a circle with radius $\varepsilon_S$ and centered at stop $j$, then

\begin{equation}
  \label{eq:l1b}
  \log L_1(\bX_k^{(i)} | \bY_k)=
  \begin{cases}
    - \log(\pi \varepsilon_S^2) & \bY_k \in B_s \\
    - \infty & \text{otherwise}
  \end{cases}
\end{equation}

%\subsubsection[L1c]{$L_1''$: at intersection}
\item
If the particle is considered to be at an intersection, we apply the following likelihood.
Similar to $L_1'$, however need to check observation is heading into the intersection
(and not out of it) by using two boundaries around the intersection and a point $\varepsilon_I$ away.
Let $B_I^1$ and $B_I^2$ be two circles with radii $\varepsilon_I$,
the first centered on the intersection and the second centered $\varepsilon_I$ meters before it, then

\begin{equation}
  \label{eq:l1c}
  \log L_1 (\bX_k^{(i)} | \bY_k) =
  \begin{cases}
    - \log(A) & \bY_k \in B_I^1 \cap B_I^2 \\
    - \infty & \text{otherwise}
  \end{cases}
\end{equation}
where $A = 2\varepsilon_I^2 \left( \cos^{-1}(\frac{1}{2}) - \frac{\sqrt{3}}{4} \right)$,
the area of $B_I^1\cap B_I^2$.

\end{enumerate}

%\subsubsection[L2]{$L_2$: not at a stop}

The second part of the likelihood involves the stop time update,
and how the likelihood is formed depends on what data is available,
where $\bT_k = \left[ j, T^a_j , T^d_j \right]^T$ 
are the stop index, arrival and departure times,
respectively, either of which can be NA.

\begin{enumerate}
\item 
  If both $T^a_j$ and $T^d_j$ are NA, 
  then there is no additional likelihood;
  $ \log L_2  (\bX_k^{(i)} | \bT_k) = 0 $.

\item
  If departure time, $T^d_j$, is available only, for stop $j$, then we
  know the bus has at least departed stop $j$; therefore,
  \begin{equation}
    \label{eq:l2}
      \log L_2 (\bX_k^{(i)} | \bT_k) =
      \begin{cases}
        - \log(2\pi\sigma_d^2) - \frac{1}{2\sigma_d^2}(T^d_j - D_j^{(i)})^2 \\
        \hfill\quad \text{if } s_k^{(i)} \geq j \text{ and $D_j^{(i)}$ is set} \\
        - \infty \\
        \hfill\quad \text{otherwise}
      \end{cases}
  \end{equation}

\item
  If arrival time, $T^a_j$, is available only, then similarly
  \begin{equation}
    \label{eq:l2b}
    \log L_2  (\bX_k^{(i)} | \bT_k) =
    \begin{cases}
      - \log(2\pi\sigma_d^2) - \frac{(T^a_j - A_j^{(i)})^2}{2\sigma_d^2} & s_k^{(i)} \geq j \\
      - \infty & \text{otherwise}
    \end{cases}.
  \end{equation}
  Note that this does not assume the particle hasn't yet left the stop;
  the departure time could simply be missing data.

\item
  If both arrival time and departure time are available, then
  \begin{equation}
    \label{eq:l2c}
    \log L_2  (\bX_k^{(i)} | \bT_k) =
    \begin{cases}
      \begin{array}
        - \log(2\pi\sigma_d^2) - \frac{(T^a_j - A_j^{(i)})^2}{2\sigma_d^2} \\
        + \log(\|\bT_k\|) - \|\bT_k\|
        (D_k^{(i)} - A_k^{(i)})
      \end{array} \\
      \hfill\quad \text{if } s_k^{(i)} \geq j \text{ and $D_j^{(i)}$ is set}  \\
      - \infty\\
      \hfill\quad \text{otherwise}
    \end{cases}
  \end{equation}
  which makes use of the observed dwell time at the stop.

\end{enumerate}
